\chapter{Atomic form factors for pair production}
\label{sec:atomic_form_factors}

The atomic form factors for electron-positron pair production by photons describe the effects due to the field of atomic electrons.
These effects are firstly the screening of the nuclear Coulomb field, described by the elastic atomic form factor $G_2^{\text{el}}$, and secondly events where the photon scatters from the electron field screened by the nucleus, described by the inelastic atomic form factor $G_2^{\text{inel}}$.
The contributions from the atomic form factors are only relevant for electron-positron pair production at small production angles which is the usual use case for high energy photon propagation.
To obtain the functions $\varphi$ and $\psi$ parameterizing the information from these atomic form factors, integrations of $G_2$ over kinematic variables are necessary.

\section{Hydrogen and Helium}

For $Z=1$, the atomic form factors are known exactly while for $Z=2$, an approximative model leading to a similar analytic expression can be used.
These form factors can be integrated analytically leading to the expressions \cite{RevModPhys.46.815}
%
\begingroup
\allowdisplaybreaks
\begin{align}
	\begin{split}
		\varphi_1 &= \frac{4}{3} \ln(Z) + 4 \ln\left(\frac{1}{2 \eta \alpha}\right) + \frac{13}{3} - 2 \ln(1 + C^2) \\&\quad- \frac{13}{2} C \arctan(C^{-1}) + \frac{1}{6}\frac{1}{1 - C^{-2}},
	\end{split}
	\\[2ex]
	\begin{split}
		\varphi_2 &= \frac{4}{3} \ln(Z) + 4 \ln\left(\frac{1}{2 \eta \alpha}\right) + \frac{11}{3} - 2 \ln(1 + C^2) \\&\quad+ 25 C^2 (1 - C \arctan(C^{-1})) - 14 C^2 \ln(1 + C^{-2}),
	\end{split}
	\\[2ex]
	\begin{split}
		\psi_1 &= \frac{8}{3} \ln(Z) + 4 \ln\left(\frac{1}{2 \eta \alpha}\right) + \frac{23}{3} - 2 \ln (1 + C^2) \\&\quad- 17.5 C \arctan{C^{-1}} + 8 C^2 \ln(1 + C^{-2}) - \frac{1}{6}\frac{1}{1 + C^{-2}},
	\end{split}
	\\[2ex]
	\begin{split}
		\psi_2 &= \frac{8}{3} \ln(Z) + 4 \ln\left(\frac{1}{2 \eta \alpha}\right) + \frac{21}{3} - 2 \ln (1 + C^2) \\&\quad- 105 C^2 (1 - C \arctan(C^{-1})) + 50 C^2 \ln(1 + C^{-2}) \\&\quad- 24 C^2 \biggl[ - \ln(C^2) \ln(1 + C^{-2}) + \Phi(1 + C^{-2}) - \Phi(1) \biggr],
	\end{split}
\end{align}
\endgroup
%
with
%
\begin{align*}
	C &= \frac{\delta}{2 \alpha m_e \eta}, & \delta &= \frac{m_e^2}{2 E_{\gamma} x (1-x)},
\end{align*}
%
where $\eta$ is defined as
%
\begin{align*}
	\eta &= 
	\begin{cases}
		1 & \text{if $Z = 1$},\\
		1.6875 & \text{if $Z = 2$},\\
	\end{cases}
\end{align*}
%
and $\Phi(x)$ is the dilogarithm function.

\section{Lithium and Beryllium}

For $Z=3$ and $Z=4$, the atomic form factors are only known numerically.
Instead of referring to these values, a simpler approximation for the form factors $G_2$ is used.
The parameters of this approximation are chosen in such a way that the results for $G_2$ are exact in the limits of complete screening and no screening.
Based on that, the functions $\varphi$, $\psi$ are \cite{RevModPhys.46.815}
%
\begingroup
\allowdisplaybreaks
\begin{align}
	\label{eqn:phi_1}
	\begin{split}
		\varphi_1 &= 2\left(1 + \ln(a^2 Z^{\sfrac{2}{3}} m_e^2)\right) - 2 \ln(1 + b^2) - 4 b \arctan(b^{-1}),
	\end{split}
	\\[2ex]
	\begin{split}
		\varphi_2 &= 2 \left(\frac{2}{3} + \ln(a^2 Z^{\sfrac{2}{3}} m_e^2 )\right) - 2 \ln(1 + b^2) \\&\quad+ 8 b^2  \biggl[ 1 - b \arctan(b^{-1}) - 0.75 \ln(1 + b^{-2}) \biggr],
	\end{split}
	\\[2ex]
	\begin{split}
		\psi_1 &= 2\left(1 + \ln(a^{\prime2} Z^{\sfrac{4}{3}} m_e^2)\right) - 2 \ln(1 + b^{\prime2}) - 4 b^{\prime} \arctan(b^{-1}),
	\end{split}
	\\[2ex]
	\label{eqn:psi_2}
	\begin{split}
		\psi_2 &= 2a \left(\frac{2}{3} + \ln(a^{\prime2} Z^{\sfrac{4}{3}} m_e^2 )\right) - 2 \ln(1 + b^{\prime2}) \\&\quad+ 8 b^{\prime2}  \biggl[ 1 - b \arctan(b^{\prime-1}) - 0.75 \ln(1 + b^{\prime-2}) \biggr],
	\end{split}
\end{align}
\endgroup
%
with
%
\begin{align*}
	a &= 
	\begin{cases}
		\frac{100}{m_e} Z^{-\sfrac{1}{3}} & \text{if $Z = 3$},\\
		\frac{106}{m_e} Z^{-\sfrac{1}{3}} & \text{if $Z = 4$},\\
	\end{cases}\\
	a^{\prime} &= 
	\begin{cases}
		\frac{418.6}{m_e} Z^{-\sfrac{2}{3}} & \text{if $Z = 3$},\\
		\frac{571.4}{m_e} Z^{-\sfrac{2}{3}} & \text{if $Z = 4$},\\
	\end{cases}
\end{align*}
and $b = a \delta$, $b^{\prime} = a^{\prime} \delta$.
%

\section{Heavier elements}

For $Z>4$, the Thomas-Fermi model of the atom is used to evaluate the atomic form factors.
The parameters from \eqref{eqn:phi_1} to \eqref{eqn:psi_2} were adapted to fit the numerical values obtained from the Thomas-Fermi model, yielding the expressions \cite{RevModPhys.46.815}
%
\begingroup
\allowdisplaybreaks
\begin{align}
	\begin{split}
		\varphi_1(\gamma) &= 20.863 - 2 \ln(1 + (0.55846 \gamma)^2) \\&\quad- 4 \biggl[ 1 - 0.6 \exp(-0.9\gamma) - 0.4 \exp(-1.5 \gamma) \biggr],
	\end{split}
	\\[2ex]
	\begin{split}
		\varphi_2(\gamma) &= \varphi_1(\gamma) - \frac{2}{3} \frac{1}{1 + 6.5\gamma + 6\gamma^2},
	\end{split}
	\\[2ex]
	\begin{split}
		\psi_1(\epsilon) &=  28.34 - 2\ln(1 + (3.621 \epsilon)^2) \\&\quad- 4 \biggl[ 1 - 0.7 \exp(- 8 \epsilon) - 0.3 \exp(-29.2\epsilon) \biggr],
	\end{split}
	\\[2ex]
	\begin{split}
		\psi_2(\epsilon) &= \psi_1(\epsilon) - \frac{2}{3} \frac{1}{1 + 40\epsilon + 400\epsilon^2},
	\end{split}
\end{align}
\endgroup
%
with
%
\begin{align*}
	\gamma &= \frac{200 \delta}{m_e Z^{\sfrac{1}{3}}}, & \epsilon &= \frac{200 \delta}{m_e Z^{\sfrac{2}{3}}}.
\end{align*}
%